### 2024-10-07 README.Rmd
updated readme

Color palettes are well organized and made consistent with each other. Moreover, they are scored on several aspects: color-blind-friendliness, the presence of intense colors (which should be avoided), the overall aesthetic harmony, and how many different hues are used. Finally, for each color palette a color for missing values is assigned, which is especially important for spatial data visualization. Currently we support several types: *categorical* (qualitative) palettes, *sequential* palettes, *diverging* palettes, and *bivariate* palettes (divided into four subtypes).

@@

@@

[<img src="vignettes/dash-3.png" width="650"/>](vignettes/dash-3.png)

What types and series are available?

```{r}
c4a_types()

c4a_series()
```

How many palettes per type x series?

```{r}
c4a_overview()
```

```{r, fig.height = 2}
c4a_plot("hcl.purple_green", 11, include.na = TRUE)
```

## Using cols4all palettes in ggplot2

@@

@@

*	Bivariate color palettes are available (besides the three main palette types: categorical, sequential, and diverging).

@@

*	Color for missing values are made explicit.

*   Palettes are made consistent with each other to enable comparison. For instance, black and white are (by default) removed from categorical palettes. Another standard that we adapt to is that all sequential palettes go from light to dark and not the other way round.

>>>>

Color palettes are well organized and made consistent with each other. Moreover, they are scored on several aspects: color-blind-friendliness, the presence of intense colors (which should be avoided), the overall aesthetic harmony, and how many different hues are used. Finally, for each color palette a color for missing values is assigned, which is especially important for spatial data visualization. Currently we support several types: *categorical* (qualitative) palettes, *sequential* palettes, *diverging* palettes, *cycling* palettes and *bivariate* palettes (divided into four subtypes).

@@

This development version will be on CRAN soon (October 2024)

@@

[<img src="vignettes/cols4all2.png" width="900"/>](vignettes/cols4all2.png)

What palettes are available? That is, by default; other palettes can be added!

```{r}
c4a_series()
```

## Using the tool

Use the tool to compare palettes and if needed analyse a palette in depth (via the other tabs).

Find a trade-off you like among the following properties (the columns in the main table):

* Colorblind friendly: Is the palette color blind friendly?
* Fair: Is the palette fair? In a fair palette, colors stand out about equally. 
* Hues: What hue ranges are used? All across the (rainbow) color spectrum, or a limited range?
* Vivid: Are there any colors that are very saturated? Perhaps a little too much?
* Contrast: Is there sufficient contrast against white, black (using WACG criteria) and between colors?
* 3D Blues: If blue is a palette color, a 3D visual illusion could appear.
* Naming (in development): How well can palette colors be named?

### Example 1

When we are looking for a fair categorical palette of seven colors that is as color blind friendly as possible, then filter on "Fair", and sort by "Colorblind-friendly":

[<img src="vignettes/cols4all_fair_sort_cbf.png" width="900"/>](vignettes/cols4all_fair_sort_cbf.png)

This gave us inspiration to develop own palettes: see these `cols4all` palettes below.

### Example 2

Say we need a diverging palette that is color blind friendly, and what to choose one by eye. Then filter by "Colorblind-friendly" and sort by "Huddle Middle L" (the hue of the left wing):

[<img src="vignettes/cols4all_div.png" width="900"/>](vignettes/cols4all_div.png)

Reverse sorting is also applied.

## Preliminary set of new `cols4all` palettes

We applied a basic heuristics to explore palettes that score well on a mix of the properties named above

[<img src="vignettes/c4a_pals.png" width="900"/>](vignettes/c4a_pals.png)

`area7`, `area8` and `area9` are fair, contain low pastel colors, and are color-blind friendly (up to 7 colors). So ideal for maps and other space-filling visualizations! These are used in [https://github.com/r-tmap/tmap](`tmap4`).

[<img src="vignettes/area7.png" width="900"/>](vignettes/area7.png)

`area7d`, `area8d` and `area9d` similar but for dark mode:.

[<img src="vignettes/area7d.png" width="900"/>](vignettes/area7d.png)

`line7`, `line8` and `line9` are colors with good contrast against both black and white, and are also colorblind-friendly to some extent. So ideal for line graphs and scatter plots:

[<img src="vignettes/line7_wt.png" width="440"/>](vignettes/line7_wt.png)
[<img src="vignettes/line7_bk.png" width="440"/>](vignettes/line7_bk.png)

Finally `friendly7` ... `friendly13` are colorblind-friendly palettes (disregarding the other properties):

[<img src="vignettes/friendly13.png" width="900"/>](vignettes/friendly13.png)

## `ggplot2` integration

@@

## Other R funtions

```{r, fig.height = 2}
c4a_plot_cvd("hcl.purple_green", 11, include.na = TRUE)
```

@@

*	Bivariate color palettes are available.

@@

*	Colors for missing values are made explicit.
